\documentclass[12pt]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Load in packages
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage[margin=1in]{geometry} % Sets 1-inch margins


%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{center}
\Large Week 1 practice problems

\medskip

\normalsize Elements of Microeconomics (discussion section 4)

\medskip

\small Jamie Hyder
\end{center}

\medskip

\subsection*{Question 1: Tradeoffs}
You are trying to decide whether to go on a beach trip during spring break. What costs would you need to consider in making this decision?

\subsection*{Question 2: PPF of a firm}
Let's start an Italian restaurant that makes pizzas and sandwiches.
\begin{enumerate}
    \item What will our production possibility frontier look like?
    \item Why will it take the shape that it has?
    \item How can we read the opportunity cost? Does it matter which part of the PPF we look at?
    \item Why might the shape change over time?
\end{enumerate}

\subsection*{Question 3: CPF of a consumer}
Now suppose we are \textit{going} to the Italian restaurant with a group of friends, and we want to decide what to order. Pizzas are \$10, sandwiches are \$5, and we have \$100 to spend.
    \begin{enumerate}
        \item What will our consumption possibility frontier look like?
        \item What is the opportunity cost of a pizza? Does it matter where on the CPF we are?
        \item What will happen to the CPF if we have \$200 to spend?
        \item What will happen to the CPF if the price of sandwiches increases to \$10?
    \end{enumerate}

\subsection*{Question 4: Italian restaurant economy}
Suppose the entire economy consists of Italian restaurants: during half of the week we work in one, and during the other half we buy food from them.
\begin{enumerate}
    \item What does the circular-flow diagram look like?
    \item What is missing from our model?
\end{enumerate}

\subsection*{Question 5: Guiding Economic Policy}
Economists are often asked to guide economic policy...
\medskip
\begin{enumerate}
    \item What are positive and normative statements?
    \begin{enumerate}
        \item Can you give an example of each?
    \end{enumerate}
    \item Why might two economists make different suggestions?
    \item Why might a politician ignore an economist's suggestions?
    
\end{enumerate}



\end{document}