\documentclass[12pt]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Load in packages
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage[top=1in, bottom=1in, left=1in, right=1in]{geometry}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{center}
\Large Chapter 4 Practice Problems

\medskip

\normalsize Elements of Microeconomics (discussion section 4)

\medskip

\small Jamie Hyder
\end{center}

\medskip

\section*{Question 1}
There is a distinction between things that shift supply/demand curves and things that cause movement along the curves.
\begin{enumerate}
    \item What factors shift a demand curve?
    \item What factors cause movement along a demand curve?
    \item What factors shift a supply curve?
    \item What factors cause movement along asupply curve?
\end{enumerate}

\section*{Question 2}
\begin{enumerate}

\item A few years ago Maryland passed a gas tax holiday, temporarily lowering the price of gasoline. Some critics said that lowering the tax would make people want to buy more gasoline and might end up actually \textit{increasing} the price.

\begin{enumerate}
    \item Will the tax decrease cause the demand curve for gasoline to shift?
    \item What are some complements and what are some substitutes for gasoline?
    \item What are some factors that might cause the demand curve for gasoline to shift?
\end{enumerate}


\item Draw supply and demand curves for the market of gasoline, and show the impact of the decrease in the gas tax.

\medskip

Does this represent a change in the demand curve or the supply curve?


\item Now suppose all cars experience a sudden increase in fuel efficiency: we can drive more miles with the same amount of gasoline. Represent this as a shift in supply or demand in our market for gasoline.


\item Now think about the two changes together; the gasoline tax is lowered, and fuel efficiency is increased. What is the net effect on the equilibrium quantity and price? Is it unambiguous?


\item Now suppose fuel efficiency suddenly gets \textit{worse}. Redo the exercise: show the impact on the demand curve, and the possible new market equilibrium when there is both a tax cut and a decrease in fuel efficiency.

\end{enumerate}

\section*{Question 3}
In the last few years, the Orioles have gone from one of the worst teams in MLB to one of the best.

\begin{enumerate}
    \item Draw the supply and demand curves for Orioles tickets.
    \item Does the supply curve look like it did in the gasoline market?
    \item Will the team's improved record effect supply or demand, and why?
    \item What will happen to equilibrium price and quantity?
\end{enumerate}

\end{document}
